%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%  THE SEARCH FOR NEUTRINOLESS DOUBLE BETA DECAY      %%%
%%%  2023 VERSION                                       %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[sn-mathphys]{sn-jnl}% Math and Physical Sciences Reference Style
%\jyear{2021}%
%\raggedbottom

\usepackage[T1]{fontenc}
\usepackage{microtype}
\usepackage{xspace}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amssymb,letterspace}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{braket}
\usepackage{bm}

\begin{document}

\input{defs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% FRONTMATTER

\title[The search for neutrinoless double beta decay]{The search for neutrinoless double beta decay}

\author[1,2]{\fnm{Juan Jos\'e} \sur{G\'omez-Cadenas}}\email{jjgomezcadenas@dipc.org}

\author[3]{\fnm{Justo} \sur{Mart\'in-Albo}}\email{justo.martin-albo@ific.uv.es}

\author[4,5]{\fnm{Javier} \sur{Men\'endez}}\email{menendez@fqa.ub.edu}

\author[6]{\fnm{Mauro} \sur{Mezzetto}}\email{mauro.mezzetto@pd.infn.it}

\author[1,2]{\fnm{Francesc} \sur{Monrabal}}\email{francesc.monrabal@dipc.org}

\author[3]{\fnm{Michel} \sur{Sorel}}\email{sorel@ific.uv.es}

\affil[1]{\orgdiv{Donostia International Physics Center}, \orgname{ERC Basque Excellence Research Centre}, \orgaddress{\city{San Sebasti\'an / Donostia}, \postcode{E-20018}, \country{Spain}}}

\affil[2]{\orgname{Ikerbasque (Basque Foundation for Science)}, \orgaddress{\city{Bilbao}, \postcode{E-48009}, \country{Spain}}}

\affil[3]{\orgdiv{Instituto de F\'isica Corpuscular (IFIC)}, \orgname{CSIC \& Universitat de Val\`encia}, \orgaddress{\city{ Paterna}, \postcode{E-46980}, \country{Spain}}}

\affil[4]{\orgdiv{Departament de F\'isica Qu\`antica i Astrof\'isica}, \orgname{Universitat de Barcelona}, \orgaddress{\city{Barcelona}, \postcode{E-08028}, \country{Spain}}}

\affil[5]{\orgdiv{Institut de Ci\`encies del Cosmos}, \orgname{Universitat de Barcelona}, \orgaddress{\city{Barcelona}, \postcode{E-08028}, \country{Spain}}}

\affil[6]{\orgdiv{Sezione di Padova}, \orgname{INFN}, \orgaddress{\city{Padova}, \postcode{I-35131}, \country{Italy}}}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\abstract{Neutrinos are the only particles in the Standard Model that could be Majorana fermions, that is, completely neutral fermions that are their own antiparticles. The most sensitive known experimental method to verify whether neutrinos are Majorana particles is the search for neutrinoless double beta decay. The last two decades have witnessed the development of a vigorous program of neutrinoless double beta decay experiments, spanning several isotopes and developing different strategies to handle the backgrounds to a hypothetical signal. In addition, remarkable progress has been made in the understanding of the neutrinoless double beta decay nuclear matrix elements, thus reducing a substantial part of the theoretical uncertainties affecting the particle physics interpretation of this process. On the other hand, the negative results by several experiments, combined with the hint that neutrino mass ordering could be normal, may imply very light neutrino masses and thus very long lifetimes for the neutrinoless double beta decay process. In this report, we review the main aspects of such process, the recent progress on theoretical ideas and the experimental state of the art. We then consider the experimental challenges to be addressed in order to increase the sensitivity to detect the process in the likely case that lifetimes are much longer than currently explored, and discuss a selection of the most promising experimental efforts.
}
\keywords{Neutrinos, Majorana, Double beta decay, Nuclear matrix elements}

\maketitle

%\tableofcontents
%\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction} \label{sec:intro} 
\input{src/intro.tex}

\section{Massive neutrinos} \label{sec:massivenu}
\input{src/massivenus.tex}

\section{Neutrinoless double beta decay} \label{sec:bb0nu}
\input{src/bb0nu.tex}

\section{Nuclear matrix elements} \label{sec:nme}
\input{src/nme.tex}

\section{Ingredients for\bbonu\ experiments} \label{sec:ingredients}
\input{src/ingredients.tex}

\section{Neutrinoless double beta decay experiments} \label{sec:experiments}
\input{src/exps.tex}

\section{Conclusions}
\input{src/conclusions.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\backmatter

\bmhead{Acknowledgments}
The authors thank Stefano Gariazzo and Lotta Jonimiemi for providing us with Figs.~\ref{fig:particlemasses} and \ref{fig:NME_density}, respectively. The authors acknowledge support from the European Research Council (ERC) under Grant Agreement No. 951281-BOLD; the European Union’s Framework Programme for Research and Innovation Horizon 2020 (2014–2020) under Grant Agreement No. 957202-HIDDEN; the MCIN/AEI of Spain and ERDF A way of making Europe under grant PID2020-118758GB-I00, grant PID2021-125475NB, ``Ram\'on y Cajal" grant RYC-2017-22781, Severo Ochoa Program grant CEX2018-000867-S and Mar\'ia de Maeztu Program grant CEX2019-000918-M; the Generalitat Valenciana of Spain under grants PROMETEO/2021/087 and CIDEGENT/2019/049. We also thank Alessandro Bettini for useful discussions and comments on the manuscript.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliography{references}

\end{document}

