%%%
The search for double beta decay has a long history, approaching now a full century. Initially, and for decades, the field was dominated by geochemical measurements (e.g., Ref.~\cite{Inghram:1950qv}), which looked for evidence of \bb-decay products in geologically old ($\sim 10^9$ years) minerals rich in the parent isotope. An excess of the daughter isotope over its natural concentration was interpreted as evidence for \bb\ decay (either \bbtnu\ or \bbonu, as the method cannot distinguish between them). 

It was not until 1987 that the \bbtnu\ decay mode was directly observed in the laboratory \cite{Elliott:1987kp,Moe:2014ioa}. The detector employed was a fairly large ($\sim1$ m$^{3}$) time projection chamber with a \bb\ source (14~g of enriched \Se{82}) deposited on a thin foil that formed the central electrode of the chamber. The trajectories of the electrons emitted from the source foil were recorded by the TPC and then analyzed to infer their energy and kinematic features. Since this initial detection, the two-neutrino mode has been directly observed for eight other isotopes in several experiments (see Table~\ref{tab:bb2nu_exp} for further details).

By contrast, no evidence of \bbonu\ decay has been found so far. Nevertheless, the progress made in the last two decades in the development of ultra-low-background detection technologies has been extraordinary. The experimental goal for the next generation of experiments is the exploration of the region of half-lives up to $10^{28}$~years. This will require exposures well beyond 1~tonne~year and background rates lower than 1~count~tonne$^{-1}$~yr$^{-1}$. Only a few of the experimental techniques presently considered will be able to attain those levels. Besides, given the scale and cost of future experiments, the consolidation of the international effort will be essential.

In this section, we review the most promising technologies for the next generation of experiments, summarizing the results of past experiments, their current status and future plans. This discussion does not pretend to be exhaustive, and the interested reader is referred to the cited publications for more detailed information.

\subsection{High-purity germanium detectors} 
\label{subsec:hpge}
\input{src/hpge}

\subsection{Bolometers} \label{subsec:bolometers}
\input{src/bolometers}

\subsection{Xenon time projection chambers} 
\label{subsec:XeTPCs}
\input{src/xetpcs}

\subsection{Loaded liquid scintillator} 
\label{subsec:liquid_scint}
\input{src/liquid_scint}


% \subsection{Past experiments} \label{subsec:past} 
% \input{src/past.tex}

% \subsection{CUORE} \label{subsec:cuore}
% \input{src/cuore.tex}

% \subsection{nEXO} \label{subsec:exo}
% \input{src/exo.tex}

% \subsection{GERDA} \label{subsec:gerda}
% \input{src/gerda.tex} 

% \subsection{MAJORANA} \label{subsec:majorana}
% \input{src/majorana.tex} 

% \subsection{KamLAND-Zen} \label{subsec:kamland}
% \input{src/kamland.tex}

% \subsection{NEXT} \label{subsec:next}
% \input{src/next.tex}

% \subsection{SNO+} \label{subsec:sno+}
% \input{src/sno+.tex}

% \subsection{SuperNEMO} \label{subsec:snemo}
% \input{src/snemo.tex}

% \subsection{Other proposals} \label{subsec:other}
% \input{src/other_exps.tex}

% \subsection{Sensitivity of new-generation experiments} \label{subsec:sensi}
% \input{src/sensi.tex}

% \subsection{Validity of sensitivity assumptions} \label{subsec:sensi_assumptions}
% \input{src/sensi_assumptions.tex}
